---
title: "week2"
author: "Carmen Pfoster"
date: "29 4 2022"
output: html_document
---

```{r}
## Load the necessary libraries ################################################

library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times

## Import the downloaded csv ##################################################

wildschwein_BE <- read_delim("wildschwein_BE_2056.txt",",") # adjust path

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)
```

```{r}
#calculating time lag
wildschwein_BE <- group_by(wildschwein_BE,TierID)
wildschwein_BE <- mutate(wildschwein_BE,timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "secs")))
```

```{r}
#some visualizations of the data
ggplot(wildschwein_BE, aes(DatetimeUTC, TierID))+
  geom_line()+
  ggtitle("measuring period per animal")

ggplot(wildschwein_BE, aes(timelag))+
  geom_histogram(breaks=seq(0, 15000, by=60))+
  scale_y_continuous(trans='log10')+
  ggtitle("Histogramm of timelags")
```

```{r}
wildschwein_BE <- wildschwein_BE %>% 
  mutate(DayMonth = format(as.Date(DatetimeUTC), "%d-%m")) 

wildschwein_BE$DayMonth <- as.Date(wildschwein_BE$DayMonth, format="%d-%m")

ggplot(wildschwein_BE, aes(DayMonth, timelag, color = TierID))+
  geom_line()+
  ggtitle("timelags per animal")+
  scale_x_date(date_labels = "%B", date_breaks = "1 month")+
  ylim(c(0, 20000))
```

####How many individuals were tracked?
There were 3 animals tracked

####For how long were the individual tracked? Are there gaps?
The tracking for animal 002A started in October 2014 and ended in July 2015. The tracking of Animal 016A started at the same time as 002A but ended at the end of June in 2015. And the animal 018A was tracked between September 2014 and July 2015.

####Were all individuals tracked concurrently or sequentially?
unfortunately I didn't manage to change the x-axis of the last graph which is why it doesn't look very intuitive. It looks like first only 018A was tracked. From November until March all 3 animals were tracked. After that the tracking was rather irregularly

####What is the temporal sampling interval between the locations?
The temporal sampling interval varies mostly between ca. 10 Minutes and 1.5 hours. There are some periods with a minimal sampling interval is 1 hour.


```{r}
wildschwein_BE <- mutate(wildschwein_BE,E2 = lead(E))
wildschwein_BE <- mutate(wildschwein_BE,N2 = lead(N))
wildschwein_BE <- mutate(wildschwein_BE, steplength = sqrt(((E-E2)^2)+(N-N2)^2))

#calculate the speed in m/s
wildschwein_BE <- mutate(wildschwein_BE, speed = steplength/timelag)
```

```{r}
caro <- read_delim("caro60.txt",",") 

caro <- st_as_sf(caro, coords = c("E", "N"), crs = 2056, remove = FALSE)
```

```{r}
seq3 <- seq(from = 1, to = nrow(caro), by = 3)
seq6 <- seq(from = 1, to = nrow(caro), by = 6)
seq9 <- seq(from = 1, to = nrow(caro), by = 9)

caro_3 <- dplyr::slice(caro, seq3)
caro_6 <- dplyr::slice(caro, seq6)
caro_9 <- dplyr::slice(caro, seq9)

nrow(caro)
nrow(caro_3)
nrow(caro_6)
nrow(caro_9)
```

```{r}
#speed caro
caro <- group_by(caro,TierID)
caro <- mutate(caro,timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "secs")),
               E2 = lead(E),
               N2 = lead(N),
               steplength = sqrt(((E-E2)^2)+(N-N2)^2),
               speed = steplength/timelag)
#speed caro_3
caro_3 <- group_by(caro_3,TierID)
caro_3 <- mutate(caro_3,timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "secs")),
               E2 = lead(E),
               N2 = lead(N),
               steplength = sqrt(((E-E2)^2)+(N-N2)^2),
               speed = steplength/timelag)

#speed caro_6
caro_6 <- group_by(caro_6,TierID)
caro_6 <- mutate(caro_6,timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "secs")),
               E2 = lead(E),
               N2 = lead(N),
               steplength = sqrt(((E-E2)^2)+(N-N2)^2),
               speed = steplength/timelag)

#speed caro_9
caro_9 <- group_by(caro_9,TierID)
caro_9 <- mutate(caro_9,timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "secs")),
               E2 = lead(E),
               N2 = lead(N),
               steplength = sqrt(((E-E2)^2)+(N-N2)^2),
               speed = steplength/timelag)
```

```{r}
ggplot(NULL, aes(E,N))+
  geom_path(data=caro, color="blue")+
  geom_path(data=caro_3, color="red")+
  ggtitle("Comparing original- with 3 minutes-resampled data")
  
ggplot(NULL, aes(E,N))+
  geom_path(data=caro, color="blue")+
  geom_path(data=caro_6, color="red")+
  ggtitle("Comparing original- with 6 minutes-resampled data")

ggplot(NULL, aes(E,N))+
  geom_path(data=caro, color="blue")+
  geom_path(data=caro_9, color="red")+
  ggtitle("Comparing original- with 9 minutes-resampled data")

ggplot(NULL, aes(DatetimeUTC, speed))+
  geom_line(data=caro, col="red")+
  geom_line(data=caro_3, col="blue")+
  geom_line(data=caro_6, col="yellow")+
  geom_line(data=caro_9, col="green")+
  ggtitle("Comparing derived speed at different sampling intervals")
```
When comparing the trajectories of the different interval samples, it can be seen, that the degree of detail reduces with increasing time intervals. Regarding the speed, the complete data set is characterized by higher speed values and more variation. Whereas the 9-minute interval is pretty flat and smooth


```{r}
library(zoo)

caro <- mutate(caro, mean_3 = rollmean(caro$speed,k = 3,fill = NA,align = "left"), mean_4 = rollmean(caro$speed,k = 4,fill = NA,align = "left"), mean_5 = rollmean(caro$speed,k = 5,fill = NA,align = "left"), mean_10 = rollmean(caro$speed,k = 10,fill = NA,align = "left"))

ggplot()+
  geom_line(data=caro,aes(DatetimeUTC, mean_3), col="red")+
  geom_line(data=caro,aes(DatetimeUTC, mean_4), col="blue")+
  geom_line(data=caro,aes(DatetimeUTC, mean_5), col="yellow")+
  geom_line(data=caro,aes(DatetimeUTC, mean_10), col="green")+
  ggtitle("Comparing derived speed at different smoothing levels")+ylab("speed [m/s]")
```
when comparing this figure to the one with the different sampling intervals one can see that applying a moving window reduces the noise but changes in speed can still be observed (different than for the different sampling intervals).
